---
title: "HumanAneurysm"
author: "A.DeMartin"
date: "2024-03-19"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r load packages, warning=FALSE, include=FALSE}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Hs.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(RColorBrewer)
```

##############################start pre-processing##############################

## load files and merge
```{r load files}
### load and merge all 
basedir <- "/Users/immbio/Desktop/Project/Angelina/HumanAneurysm/data/AneurysmDataMarch2024/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratM")){
    seuratM <- merge(x = seuratM, y = seuratS)
  }else{
    seuratM <- seuratS
  }
}

remove(seuratS)
table(seuratM$dataset)
table(seuratM$orig.ident)

### load and merge all 
basedir <- "/Users/immbio/Desktop/Project/Angelina/HumanAneurysm/data/AneurysmDataApril2024/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratM")){
    seuratM <- merge(x = seuratM, y = seuratS)
  }else{
    seuratM <- seuratS
  }
}

remove(seuratS)
table(seuratM$dataset)
table(seuratM$orig.ident)

### load and merge all 
basedir <- "/Users/immbio/Desktop/Project/Angelina/HumanAneurysm/data/AneurysmDataJuly2024/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratM")){
    seuratM <- merge(x = seuratM, y = seuratS)
  }else{
    seuratM <- seuratS
  }
}

remove(seuratS)
table(seuratM$dataset)
table(seuratM$orig.ident)
```

## add metadata
```{r add metadata}
### add any type of metadata
### patient
control_01 <- c("344261_07-7_20240219_Hu_nucseq_control_01")
UIA_04 <- c("344261_08-8_20240219_Hu_nucseq_UIA_04")
UIA_07 <- c("344261_09-9_20240219_Hu_nucseq_UIA_07")
tArm_ctrl_1 <- c("348751_2-7_20240326_Hu_nucseq_tArm_ctrl_1")
tArm_ctrl_2 <- c("348751_3-8_20240326_Hu_nucseq_tArm_ctrl_2")
UIA_001 <- c("348751_4-9_20240326_Hu_nucseq_UIA_001")
UIA_3 <- c("348751_5-10_20240326_Hu_nucseq_UIA_003")
UIA_5 <- c("348751_6-11_20240326_Hu_nucseq_UIA_005")
UIA_012 <- c("348751_7-12_20240326_Hu_nucseq_UIA_012")
UIA_002 <- c("348751_8-13_20240326_Hu_nucseq_UIA_002")
Cadaver_ctrl_01 <- c("353921_25-25_20240529_Hu_nucseq_Cadaver_ctrl_01")
STA_LG <- c("353921_26-26_20240529_Hu_nucseq_STA_LG")
Ocp_ctrl_03 <- c("353921_27-27_20240522_Hu_nucseq_Ocp_ctrl_03")

seuratM$patient <- "pat_nr"
seuratM$patient[which(seuratM$dataset %in% control_01)] <- "control_01"
seuratM$patient[which(seuratM$dataset %in% UIA_04)] <- "UIA_04"
seuratM$patient[which(seuratM$dataset %in% UIA_07)] <- "UIA_07"
seuratM$patient[which(seuratM$dataset %in% tArm_ctrl_1)] <- "tArm_ctrl_1"
seuratM$patient[which(seuratM$dataset %in% tArm_ctrl_2)] <- "tArm_ctrl_2"
seuratM$patient[which(seuratM$dataset %in% UIA_001)] <- "UIA_001"
seuratM$patient[which(seuratM$dataset %in% UIA_3)] <- "UIA_3"
seuratM$patient[which(seuratM$dataset %in% UIA_5)] <- "UIA_5"
seuratM$patient[which(seuratM$dataset %in% UIA_012)] <- "UIA_012"
seuratM$patient[which(seuratM$dataset %in% UIA_002)] <- "UIA_002"
seuratM$patient[which(seuratM$dataset %in% Cadaver_ctrl_01)] <- "Cadaver_ctrl_01"
seuratM$patient[which(seuratM$dataset %in% STA_LG)] <- "STA_LG"
seuratM$patient[which(seuratM$dataset %in% Ocp_ctrl_03)] <- "Ocp_ctrl_03"
table(seuratM$patient)

### batch
batchMarch24 <- c("344261_07-7_20240219_Hu_nucseq_control_01","344261_08-8_20240219_Hu_nucseq_UIA_04","344261_09-9_20240219_Hu_nucseq_UIA_07")
batchApril24 <- c("348751_2-7_20240326_Hu_nucseq_tArm_ctrl_1","348751_3-8_20240326_Hu_nucseq_tArm_ctrl_2","348751_4-9_20240326_Hu_nucseq_UIA_001", "348751_5-10_20240326_Hu_nucseq_UIA_003", "348751_6-11_20240326_Hu_nucseq_UIA_005", "348751_7-12_20240326_Hu_nucseq_UIA_012", "348751_8-13_20240326_Hu_nucseq_UIA_002")
batchJuly24 <- c("353921_25-25_20240529_Hu_nucseq_Cadaver_ctrl_01","353921_26-26_20240529_Hu_nucseq_STA_LG","353921_27-27_20240522_Hu_nucseq_Ocp_ctrl_03")

seuratM$batch <- "batch"
seuratM$batch[which(seuratM$dataset %in% batchMarch24)] <- "batchMarch24"
seuratM$batch[which(seuratM$dataset %in% batchApril24)] <- "batchApril24"
seuratM$batch[which(seuratM$dataset %in% batchJuly24)] <- "batchJuly24"

table(seuratM$batch)
```

##set color vectors 
```{r set color vector}
col <- c("#202547","#BE3144","#727077","#355C7D","#779d8d","#dfc27d","#f4a582","#B45B5C")
names(col) <- c("0", "1", "2", "3", "4", "5", "6")

#get colors
#brewer.pal(n= 10, name = "Spectral")
#"#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B" "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
#brewer.pal(n= 10, name = "Accent")
#"#7FC97F" "#BEAED4" "#FDC086" "#FFFF99" "#386CB0" "#F0027F" "#BF5B17" "#666666"

colpat <- c("#9E0142", "#D53E4F", "#F46D43", "#FDAE61", "#FEE08B", "#E6F598", "#ABDDA4", "#66C2A5", "#3288BD", "#5E4FA2" ,"#BEAED4" , "#FDC086", "#727077")
names(colpat) <- unique(seuratM$patient)

colbatch <- c("#779d8d","#dfc27d","#f4a582")
names(colbatch) <- unique(seuratM$batch)
```

## QC merged 
```{r QC merged}
### QC merged
# Extract meta.data from the Seurat object
meta.data <- seuratM@meta.data
# Create the density plot
ptotalpat <- ggplot(data = meta.data, aes(x = total, color = patient, fill = patient)) +
  geom_density(alpha = 0.2) +
  scale_fill_manual(values = colpat) +
  scale_color_manual(values = colpat) +
  theme_classic() +
  scale_x_log10() +
  ylab("density") +
  geom_vline(xintercept = 100)

pdetectedpat <- ggplot(data = meta.data, aes(x = detected, color = patient, fill = patient)) +
  geom_density(alpha = 0.2) +
  scale_fill_manual(values = colpat) +
  scale_color_manual(values = colpat) +
  theme_classic() +
  scale_x_log10() +
  ylab("density") +
  geom_vline(xintercept = 100)

ptotalbatch <- ggplot(data = meta.data, aes(x = total, color = batch, fill = batch)) +
  geom_density(alpha = 0.2) +
  scale_fill_manual(values = colbatch) +
  scale_color_manual(values = colbatch) +
  theme_classic() +
  scale_x_log10() +
  ylab("density") +
  geom_vline(xintercept = 100)

pdetectedbatch <- ggplot(data = meta.data, aes(x = detected, color = batch, fill = batch)) +
  geom_density(alpha = 0.2) +
  scale_fill_manual(values = colbatch) +
  scale_color_manual(values = colbatch) +
  theme_classic() +
  scale_x_log10() +
  ylab("density") +
  geom_vline(xintercept = 100)
# Return the plots as a list
list(ptotalpat, pdetectedpat, ptotalbatch ,pdetectedbatch)
```
## filter and rerun seurat
```{r rerun seurat, eval=FALSE, include=TRUE}
##filter for Cadaver_ctr_01, UIA_07, UIA_04, Ocp_ctrl_03, control_01, tArm_ctr_1
seuratMfil <- subset(seuratM, patient %in% c("Cadaver_ctrl_01", "UIA_07", "UIA_04", "Ocp_ctrl_03", "control_01", "tArm_ctrl_1"))
table(seuratMfil$dataset)

seuratM <- seuratMfil
remove(seuratMfil)
table(seuratM$dataset)

#rerun seurat
seuratM <- NormalizeData (object = seuratM)
seuratM <- FindVariableFeatures(object = seuratM)
seuratM <- ScaleData(object = seuratM, verbose = TRUE)
seuratM <- RunPCA(object=seuratM, npcs = 30, verbose = FALSE)
seuratM <- RunTSNE(object=seuratM, reduction="pca", dims = 1:20)
seuratM <- RunUMAP(object=seuratM, reduction="pca", dims = 1:20)
seuratM <- FindNeighbors(object = seuratM, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratM <- FindClusters(object = seuratM, resolution = res[i], random.seed = 1234)
}
```

```{r save merged seurat object, eval=FALSE, include=TRUE}
### save seurat object
saveRDS(seuratM, file="/Users/immbio/Desktop/Project/Angelina/HumanAneurysm/data/Aneuryms_allmerged_seurat.rds")
```

##############################end pre-processing##############################

## load file
```{r load merged file}
##load merged file 
fileNam <- "/Users/immbio/Desktop/Project/Angelina/HumanAneurysm/data/Aneuryms_allmerged_seurat.rds"
seuratM <- readRDS(fileNam)

colpat2 <- c("#202547","#BE3144","#727077","#779d8d","#dfc27d","#f4a582")
names(colpat2) <- unique(seuratM$patient)

table(seuratM$dataset)
table(seuratM$RNA_snn_res.0.25)
table(seuratM$orig.ident)
```

## plot cell numbers
```{r cellnumber circle}
cell_count <- data.frame(table(seuratM$patient))
colnames(cell_count) <- c("patient", "Freq")
hsize <- 1.5

ggplot(cell_count, aes(x = hsize, y = Freq, fill = patient)) +
  scale_fill_manual(values = colpat2) +
  geom_col(color = "white") +
  coord_polar(theta = "y") +
  xlim(c(0.2, hsize + 0.5)) +
  theme_void() +
  ggtitle("cell number") +
  theme(plot.title = element_text(hjust = 0.5, size = 15)) +
  geom_text(aes(label = Freq), position = position_stack(vjust = 0.5))
```

## plot umaps
```{r umap}
Idents(seuratM) <- seuratM$RNA_snn_res.0.25
DimPlot(seuratM, reduction = "umap", pt.size = 0.8) 

Idents(seuratM) <- seuratM$patient
DimPlot(seuratM, reduction = "umap", pt.size = 0.8, cols = colpat2) 
```

## calculate marker genes
```{r marker genes, include=TRUE, eval=FALSE}
##cluster marker
Idents(seuratM) <- seuratM$RNA_snn_res.0.25
markerGenes <- FindAllMarkers(seuratM, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01)
#save table
write.table(markerGenes, 
            file= "/Users/immbio/Desktop/Project/Angelina/HumanAneurysm/analysis/markerGenes",
            sep="\t",
            quote=F,
            row.names=F,
            col.names=T)
```

## plot features
```{r feature plot}
FeaturePlot(seuratM, features = "ENSG00000133392.MYH11", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000107796.ACTA2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000076706.MCAM", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000249669.CARMN", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000168685.IL7R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000116824.CD2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000110848.CD69", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000081237.PTPRC", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000154258.ABCA9", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000159167.STC1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000196569.LAMA2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000047457.CP", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000134853.PDGFRA", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000081052.COL4A4", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000164530.PI16", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000000971.CFH", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000177575.CD163", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000018280.SLC11A1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000164733.CTSB", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000010327.STAB1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000261371.PECAM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000172889.EGFL7", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000110799.VWF", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000107147.KCNT1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000168824.NSG1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000164188.RANBP3L", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000158296.SLC13A3", pt.size = 1, cols = c("lightgrey", "#BE3144"))
	
```
######################################## integrate over batches ########################################################

## integrate over batches
```{r integrate data}
seuratM.list <- SplitObject(object = seuratM, split.by = "batch")
for (i in 1:length(x = seuratM.list)) {
    seuratM.list[[i]] <- NormalizeData(object = seuratM.list[[i]],
                                      verbose = FALSE)
    seuratM.list[[i]] <- FindVariableFeatures(object = seuratM.list[[i]], 
        selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}

seuratM.anchors <- FindIntegrationAnchors(object.list = seuratM.list, dims = 1:10,
                                         k.filter = 200)
seuratM.int <- IntegrateData(anchorset = seuratM.anchors, dims = 1:10)
DefaultAssay(object = seuratM.int) <- "integrated"

# rerun seurat
seuratM.int <- ScaleData(object = seuratM.int, verbose = FALSE,
                        features = rownames(seuratM.int))
seuratM.int <- RunPCA(object = seuratM.int, npcs = 20, verbose = FALSE)
#seuratM.int <- RunTSNE(object = seuratM.int, reduction = "pca", dims = 1:10)
seuratM.int <- RunUMAP(object = seuratM.int, reduction = "pca", dims = 1:10)

seuratM.int <- FindNeighbors(object = seuratM.int, reduction = "pca", dims = 1:10)
res <- c(0.25,0.6,0.4,0.8)
for(i in 1:length(res)){
  seuratM.int <- FindClusters(object = seuratM.int, resolution = res[i],
                             random.seed = 1234)
}

DefaultAssay(object = seuratM.int) <- "RNA"
```

## visualize data integrated over batches
```{r visualize int cluster over batches}
Idents(seuratM.int) <- seuratM.int$integrated_snn_res.0.25
DimPlot(seuratM.int, reduction="umap", pt.size = 1)

Idents(seuratM.int) <- seuratM.int$batch
DimPlot(seuratM.int, reduction="umap", cols = colbatch, pt.size = 1)

Idents(seuratM.int) <- seuratM.int$patient
DimPlot(seuratM.int, reduction="umap", cols = colpat2, pt.size = 1)
```
######################################## end integrate over batches ########################################################

######################################## integrate over patients ########################################################

## integrate over patients
```{r integrate data over patients}
seuratM.list <- SplitObject(object = seuratM, split.by = "patient")
for (i in 1:length(x = seuratM.list)) {
    seuratM.list[[i]] <- NormalizeData(object = seuratM.list[[i]],
                                      verbose = FALSE)
    seuratM.list[[i]] <- FindVariableFeatures(object = seuratM.list[[i]], 
        selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}

seuratM.anchors <- FindIntegrationAnchors(object.list = seuratM.list, dims = 1:10,
                                         k.filter = 200)
seuratM.intpat <- IntegrateData(anchorset = seuratM.anchors, dims = 1:10)
DefaultAssay(object = seuratM.intpat) <- "integrated"

# rerun seurat
seuratM.intpat <- ScaleData(object = seuratM.intpat, verbose = FALSE,
                        features = rownames(seuratM.intpat))
seuratM.intpat <- RunPCA(object = seuratM.intpat, npcs = 20, verbose = FALSE)
#seuratM.int <- RunTSNE(object = seuratM.int, reduction = "pca", dims = 1:10)
seuratM.intpat <- RunUMAP(object = seuratM.intpat, reduction = "pca", dims = 1:10)

seuratM.intpat <- FindNeighbors(object = seuratM.intpat, reduction = "pca", dims = 1:10)
res <- c(0.25,0.6,0.4,0.8)
for(i in 1:length(res)){
  seuratM.intpat <- FindClusters(object = seuratM.intpat, resolution = res[i],
                             random.seed = 1234)
}
DefaultAssay(object = seuratM.intpat) <- "RNA"
```

## visualize data integrated over patients
```{r visualize int cluster over patients}
#set color vector
col <- c("#727077","#BE3144","#355C7D","#779d8d","#f4a582","#B45B5C","#dfc27d")
names(col) <- c("0", "1", "2", "3", "4", "5", "6")

Idents(seuratM.intpat) <- seuratM.intpat$integrated_snn_res.0.25
DimPlot(seuratM.intpat, reduction="umap", cols = col, pt.size = 1)
DimPlot(seuratM.intpat, reduction="umap", cols = col, pt.size = 1) + theme(legend.position = "null")

Idents(seuratM.intpat) <- seuratM.intpat$batch
DimPlot(seuratM.intpat, reduction="umap", cols = colbatch, pt.size = 1)
DimPlot(seuratM.intpat, reduction="umap", cols = colbatch, pt.size = 1) + theme(legend.position = "null")

Idents(seuratM.intpat) <- seuratM.intpat$patient
DimPlot(seuratM.intpat, reduction="umap", cols = colpat2, pt.size = 1)
DimPlot(seuratM.intpat, reduction="umap", cols = colpat2, pt.size = 1) + theme(legend.position = "null")
```
## calculate marker genes integrated over patients
```{r marker genes intpat, include=TRUE, eval=FALSE}
##cluster marker
Idents(seuratM.intpat) <- seuratM.intpat$integrated_snn_res.0.25
markerGenes <- FindAllMarkers(seuratM, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01)
#save table
write.table(markerGenes, 
            file= "/Users/immbio/Desktop/Project/Angelina/HumanAneurysm/analysis/markerGenesintpat",
            sep="\t",
            quote=F,
            row.names=F,
            col.names=T)
```

## plot features intpat
```{r feature plot intpat}
FeaturePlot(seuratM.intpat, features = "ENSG00000133392.MYH11", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000107796.ACTA2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000076706.MCAM", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000249669.CARMN", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM.intpat, features = "ENSG00000168685.IL7R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000116824.CD2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000110848.CD69", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000081237.PTPRC", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM.intpat, features = "ENSG00000154258.ABCA9", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000159167.STC1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000196569.LAMA2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000047457.CP", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000134853.PDGFRA", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000081052.COL4A4", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000164530.PI16", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000000971.CFH", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM.intpat, features = "ENSG00000177575.CD163", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000018280.SLC11A1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000164733.CTSB", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000010327.STAB1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM.intpat, features = "ENSG00000261371.PECAM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000172889.EGFL7", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000110799.VWF", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM.intpat, features = "ENSG00000107147.KCNT1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000168824.NSG1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000164188.RANBP3L", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM.intpat, features = "ENSG00000158296.SLC13A3", pt.size = 1, cols = c("lightgrey", "#BE3144"))
```

## abundance plots
```{r abundance plot, fig.height=8, fig.width=10}
###dataset
datList <- NULL
for(con in unique(seuratM.intpat$dataset)){
  seuratSub <- subset(seuratM.intpat, dataset==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$integrated_snn_res.0.25)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(dataset=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

## plot abundance
ggbarplot(dat_all, x= "dataset", y= "percent", fill = "Var1", palette = col, legend = "right", legend.titel = "cluster", ylab = "frequency")  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
######################################## end integrate over patients ########################################################

## plot umap not integrated and color by integrated clustering
```{r umap colored by intpat clustering}
Idents(seuratM) <- seuratM.int$integrated_snn_res.0.25
DimPlot(seuratM, reduction = "umap", pt.size = 1) 

Idents(seuratM) <- seuratM.intpat$integrated_snn_res.0.25
DimPlot(seuratM, reduction = "umap", pt.size = 1) 
```

## session info
```{r date and session info}
date()
sessionInfo()
```
